\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-11-11 tex/ms.tex
more text

Since the Wigner matrices are block diagonal, we may evaluate the moments of the
distribution one spherical harmonic degree at a time. To that end, let us
write

@@

\begin{proof}{test_ephi}
    \left(
    \mathbf{e}_\phi^0
    \,\,\,
    \mathbf{e}_\phi^1
    \,\,\,
    \mathbf{e}_\phi^2
    \,\,\,
    \cdots
    \,\,\,
    \mathbf{e}_\phi^{l_{\mathrm{max}}}
    \right)^\top

@@

\begin{proof}{test_ephi}
    \mathbf{e}_\phi^l

@@

    \mathbf{e}^l_r \,

@@

    \pmb{\rho}^l

@@

    \mathbf{e}^l_r \,
    \cdots
\xxx{work in progress...}
In the expressions above, we define the vectors

@@

    \pmb{\rho}^l

@@

    \bar{\mathbf{e}}^l_r \,

@@

    \bar{\mathbf{e}}^l_r

@@

    \mathbf{e}^l_r

@@

We may write the terms of $\pmb{\rho}^l$ as

@@

    \rho^l_m

@@

    \sum\limits_{m'}
    {D^l_{\hat{\mathbf{x}}}}_{m,m'}(\phi) \,
    {\bar{e}^l}{{}_r}_{m'} \,

@@

    \sum\limits_{m'}
    {\bar{e}^l}{{}_r}_{m'}
    \mathrm{e}^{\frac{\imag \pi}{2}(m - m')}
    \sum\limits_i c_{m,m',i}^{l}

@@

    q^l_i(\pmb{\theta}_{\phi})

@@

@@

\begin{proof}{test_qli}
    q^l_i(\pmb{\theta}_{\phi})

@@

    \mathrm{sgn}(\sin\phi)^{i}

@@

    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
    (1 + \cos\phi)^\frac{i}{2}

@@

        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
        (1 + x)^\frac{i}{2}

@@

         & i \,\, \mathrm{even}

@@

         & i \,\, \mathrm{odd} \quad,

@@

and in the last line we made use of the transformation $x = \cos\phi$.

@@

The integral in the expression above has a closed-form solution in terms
of the hypergeometric function $_2F_1$:

@@

\begin{proof}{test_qli}
    \label{eq:qli}
    q^l_i(\pmb{\theta}_{\phi})

@@

                \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}

@@

                -\frac{i}{2},

@@

                l + \alpha + \beta - \frac{i}{2};

@@

                 & i \,\, \mathrm{odd} \quad.

@@

In order to compute the integral $\mathbf{e}_\phi$
(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qli})
for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
which can be done efficiently via upward recursion relations for both
the gamma function and the hypergeometric function.

>>>>

\subsubsection{First moment}
%
Since the Wigner matrices are block diagonal, we may evaluate the moments of the
distribution one spherical harmonic degree at a time. To that end, let us
write the first moment integral as
%
\begin{proof}{test_ephi}
    \mathbf{e}_\phi
    & \equiv
    \int
    \mathbf{R}_{\hat{\mathbf{x}}}(\phi) \,
    \mathbf{e}_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \nonumber
    \\
    %
    & =
    %
    \left(
    \mathbf{e}_\phi^0
    \,\,\,
    \mathbf{e}_\phi^1
    \,\,\,
    \mathbf{e}_\phi^2
    \,\,\,
    \cdots
    \,\,\,
    \mathbf{e}_\phi^{l_{\mathrm{max}}}
    \right)^\top
    \quad,
\end{proof}
%
where
%
\begin{proof}{test_ephi}
    \mathbf{e}_\phi^l
    & =
    \int
    \mathbf{R}^l_{\hat{\mathbf{x}}}(\phi) \,
    \mathbf{e}^l_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \nonumber \\
    %
    & =
    %
    {\mathbf{U}^l}^{-1}
    \mathbf{p}^l
    \quad,
\end{proof}
%
and we define
%
\begin{align}
    \label{eq:pl}
    \mathbf{p}^l
     & \equiv
    \int
    \mathbf{D}^l_{\hat{\mathbf{x}}}(\phi) \,
    \bar{\mathbf{e}}^l_r \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    \\
    \bar{\mathbf{e}}^l_r
     & \equiv
    \mathbf{U}^l
    \mathbf{e}^l_r
    \quad.
\end{align}
%
The integral $\mathbf{p}^l$ defined above has a closed-form solution.
To show this, we write the terms of $\mathbf{p}^l$ as
%
\begin{align}
    p^l_m
     & =
    \int
    \sum\limits_{\mu=-l}^l
    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
    {\bar{e}^l}{{}_r}_{\mu} \,
    p(\phi \, \big| \, \pmb{\theta}_{\phi}) \,
    \mathrm{d}\phi
    %
    \nonumber \\[0.5em]
    %
     & =
    \dfrac{\Gamma(\alpha + \beta)}{\Gamma(\alpha)\Gamma(\beta)}
    \sum\limits_{\mu=-l}^l
    {\bar{e}^l}{{}_r}_{\mu}
    \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu)}
    \sum\limits_{i=0}^{2l} c_{m,\mu,i}^{l}
    \,
    q^l_i(\pmb{\theta}_{\phi})
    \quad,
\end{align}
%
where
%
\begin{proof}{test_qli}
    q^l_i(\pmb{\theta}_{\phi})
    & \equiv
    \frac{1}{2}
    \int_{-\frac{\pi}{2}}^{\frac{\pi}{2}}
    %
    \mathrm{sgn}(\sin\phi)^{i}
    \big| \sin\phi \big|
    (\cos\phi)^{\alpha - 1}
    (1 - \cos\phi)^{l + \beta - \frac{i}{2} - 1}
    (1 + \cos\phi)^\frac{i}{2}
    %
    \,
    \mathrm{d}\phi
    \nonumber \\[0.5em]
    %
    & =
    %
    \begin{cases}
        \displaystyle\int_{0}^{1}
        %
        x^{\alpha - 1}
        (1 - x)^{l + \beta - \frac{i}{2} - 1 }
        (1 + x)^\frac{i}{2}
        %
        \,
        \mathrm{d}x
        %
         & i \,\, \mathrm{even}
        %
        \\
        %
        0
        %
         & i \,\, \mathrm{odd} \quad,
    \end{cases}
\end{proof}
%
and in the last line we made use of the transformation $x = \cos\phi$.
%
The integral in the expression above has a closed-form solution in terms
of the hypergeometric function $_2F_1$:
%
\begin{proof}{test_qli}
    %
    \label{eq:qli}
    q^l_i(\pmb{\theta}_{\phi})
    & =
    \resizebox{.75\hsize}{!}{$
            \begin{cases}
                \dfrac{\Gamma(\alpha)\Gamma(l + \beta - \frac{i}{2})}{\Gamma(l + \alpha + \beta - \frac{i}{2})}
                \,
                _2F_1\Big(
                -\frac{i}{2},
                \alpha;
                l + \alpha + \beta - \frac{i}{2};
                -1
                \Big)
                %
                 & i \,\, \mathrm{even}
                %
                \\
                %
                0
                %
                 & i \,\, \mathrm{odd} \quad.
            \end{cases}
        $}
\end{proof}
%
In order to compute the integral $\mathbf{e}_\phi$
(Equation~\ref{eq:e2}) we must evaluate Equation~(\ref{eq:qli})
for all $0 \leq l \leq l_{\mathrm{max}}$, $0 \leq i \leq 2l$,
which can be done efficiently via upward recursion relations for both
the gamma function and the hypergeometric function.

\subsubsection{Second moment}

@@

Similarly as before, let us write the second moment integral as

@@

@@

    \mathbf{E}^l_r \,
    {\mathbf{E}^l_r}^\top \,

@@

    \mathbf{P}^l
    {{\mathbf{U}^l}^{-1}}^\top

@@

and we define

@@

    \label{eq:Pl}
    \mathbf{P}^l

@@

    \bar{\mathbf{E}}^l_r \,
    {\mathbf{D}^l_{\hat{\mathbf{x}}}}^\top(\phi) \,

@@

    \bar{\mathbf{E}}^l_r

@@

    \mathbf{E}^l_r
    {\mathbf{E}^l_r}^\top
    {\mathbf{U}^l}^\top

@@

As before, we may express the solution to the integral $\mathbf{P}^l$ in
closed form. Let us write the terms of $\mathbf{P}^l$ as

@@

    P^l_{m,m'}

@@

    \sum\limits_{\mu=-l}^l
    \sum\limits_{\nu=-l}^l
    {D^l_{\hat{\mathbf{x}}}}_{m,\mu}(\phi) \,
    {\bar{E}^l}{{}_r}_{\mu,\nu} \,
    {D^l_{\hat{\mathbf{x}}}}_{\nu,m'}(\phi) \,

@@

        \sum\limits_{\mu=-l}^l
        \sum\limits_{\nu=-l}^l
        {\bar{E}^l}{{}_r}_{\mu,\nu}
        \mathrm{e}^{\frac{\imag \pi}{2}(m - \mu + \nu - m')}
        \sum\limits_{i=0}^{2l}
        \sum\limits_{j=0}^{2l}
        c_{m,\mu,i}^{l}
        c_{\nu,m',j}^{l}

@@

        Q^l_{i,j}(\pmb{\theta}_{\phi})

@@

where, similarly to before,

@@

    Q^l_{i,j}(\pmb{\theta}_{\phi})

@@

    \mathrm{sgn}(\sin\phi)^{i+j}

@@

    (1 - \cos\phi)^{2l + \beta - \frac{i+j}{2} - 1}
    (1 + \cos\phi)^\frac{i+j}{2}

@@

        (1 - x)^{2l + \beta - \frac{i+j}{2} - 1 }
        (1 + x)^\frac{i+j}{2}

@@

         & i+j \,\, \mathrm{even}

@@

         & i+j \,\, \mathrm{odd} \quad.

@@

We may again express this integral in closed form:

@@

@@

    \label{eq:Qlij}
    Q^l_{i,j}(\pmb{\theta}_{\phi})

@@

                \dfrac{\Gamma(\alpha)\Gamma(2l + \beta - \frac{i+j}{2})}{\Gamma(2l + \alpha + \beta - \frac{i+j}{2})}

@@

                -\frac{i+j}{2},

@@

                2l + \alpha + \beta - \frac{i+j}{2};

@@

                 & i+j \,\, \mathrm{odd} \quad.

@@

As before, this integral may be evaluated recursively to efficiently compute all
of the terms in $\mathbf{E}_\phi$.
